---
title: "Getting started with didimpute"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with didimpute}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

**didimpute** implements the imputation estimator of Borusyak, Jaravel and Spiess (2024, *Review of Economic Studies*) for staggered difference-in-differences (DiD) designs.  The estimator proceeds in two steps:

1. Fit a two-way fixed effects (TWFE) model on *untreated* observations to obtain a counterfactual prediction $\hat{y}_{it}(0)$ for each treated observation.
2. Aggregate the unit-level treatment effects $\hat{\tau}_{it} = y_{it} - \hat{y}_{it}(0)$ into researcher-specified estimands (average treatment effect on the treated, event-study coefficients, etc.).

Standard errors are cluster-robust and use the smartweights influence-function approach of the original Python package.

## Installation

```r
remotes::install_github("xiangao/didimpute")
```

## Data format

The estimator requires a long-format panel with:

- `i`: unit identifier
- `t`: time variable (integer-valued)
- `Ei`: first treated period (`NA` for never-treated units)
- `y`: outcome variable

```{r data}
library(didimpute)

# Synthetic staggered panel: 30 units, 10 periods
set.seed(42)
n_units <- 30; n_t <- 10

panel <- expand.grid(i = seq_len(n_units), t = seq_len(n_t))

# Three treatment cohorts: early (t=5), late (t=7), never-treated
panel$Ei <- ifelse(panel$i <= 10, 5L,
              ifelse(panel$i <= 20, 7L, NA_integer_))

# DGP: true effect = 1.5, with cohort heterogeneity
panel$y <- 1.5 * (!is.na(panel$Ei) & panel$t >= panel$Ei) +
             0.5 * (panel$i <= 10) * (!is.na(panel$Ei) & panel$t >= panel$Ei) +
             rnorm(nrow(panel), sd = 0.5)
```

## Baseline estimate

The simplest call returns the average treatment effect on the treated (ATT) pooled across all treated observations.

```{r baseline}
res_ate <- did_impute(panel, y = "y", i = "i", t = "t", Ei = "Ei")
print(res_ate)
```

## Event-study estimates

Pass `horizons` to estimate separate effects at each relative-time horizon.  Use `pretrends` to also estimate placebo pre-trend coefficients.

```{r eventstudy}
res <- did_impute(
  panel,
  y         = "y",
  i         = "i",
  t         = "t",
  Ei        = "Ei",
  horizons  = 0:2,
  pretrends = 2
)

summary(res)
```

The summary returns a tidy `data.frame` with one row per estimand: post-treatment effects (`tau0`, `tau1`, `tau2`) followed by pre-trend placebo coefficients (`pre1`, `pre2`).

## Event-study plot

`event_plot()` produces a `ggplot2` figure with pre-trends on the left and post-treatment effects on the right.

```{r plot, fig.width=6, fig.height=4}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  event_plot(res)
}
```

The `rarea` variant draws a shaded confidence band instead of error bars:

```{r plot-rarea, fig.width=6, fig.height=4}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  event_plot(res, plot_type = "rarea")
}
```

## Additional options

### Horizon-balanced sample (`hbalance`)

When `hbalance = TRUE`, the estimator restricts to cohorts observed at *every* requested horizon, ensuring a balanced event-study sample.

```{r hbalance}
res_bal <- did_impute(
  panel,
  y        = "y",
  i        = "i",
  t        = "t",
  Ei       = "Ei",
  horizons = 0:2,
  hbalance = TRUE
)
summary(res_bal)
```

### All observed horizons (`allhorizons`)

```{r allhorizons}
res_all <- did_impute(
  panel,
  y           = "y",
  i           = "i",
  t           = "t",
  Ei          = "Ei",
  allhorizons = TRUE
)
summary(res_all)
```

### Analytic weights

Pass the name of a weight column to `aw`:

```{r weights}
panel$pop <- runif(nrow(panel), 0.5, 1.5)  # synthetic population weight
res_w <- did_impute(
  panel,
  y  = "y",
  i  = "i",
  t  = "t",
  Ei = "Ei",
  aw = "pop"
)
print(res_w)
```

### Skipping standard errors

For large panels, passing `nose = TRUE` skips the (expensive) influence-function SE computation.

```{r nose}
res_nose <- did_impute(
  panel,
  y    = "y",
  i    = "i",
  t    = "t",
  Ei   = "Ei",
  nose = TRUE
)
print(res_nose)
```

## The `V` field

The returned object includes a square covariance matrix `V` of dimension $K \times K$, where $K$ is the total number of reported estimands (effects, then pre-trends, then controls, in that order).  Row and column names match the estimand names.  `sqrt(diag(res$V))` reproduces the reported per-estimand standard errors to numerical precision.  This is a deliberate improvement over the upstream Python package, whose `V` was a scalar (sum of squared standard errors).  `V` is `NULL` when `nose = TRUE` or when no standard-error components are present.

## Unimplemented options

The following arguments match the Python package API but are not yet implemented and will raise an error if supplied: `timecontrols`, `unitcontrols`, `leaveoneout`, `hetby`, `project`.

## Reference

Borusyak, K., Jaravel, X., and Spiess, J. (2024). Revisiting Event-Study Designs: Robust and Efficient Estimation. *Review of Economic Studies*, 91(6), 3253--3285.
